---
title: "Chlorination Cessation Alters Greenhouse Gas Dynamics in Artificial Urban Ponds"
author: "Jorge Montes Pérez"
date: "2025-03-05"
output:
  html_document:
    theme: simplex
    toc: true
    toc_depth: 4
    toc_float: 
      collapsed: false
      smooth_scroll: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

# Introduction

This is the code used to generated results and plots presented in **Chlorination Cessation Alters Greenhouse Gas Dynamics in Artificial Urban Ponds - Montes-Pérez et al., 2025** research article.

# Download data from EDI repository

```{r , echo=TRUE, results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
options(HTTPUserAgent="EDI_CodeGen")
	      

inUrl1  <- "https://pasta-s.lternet.edu/package/data/eml/edi/1927/2/3e128569a1bbc4f38cc88bd46e8a310b" 
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

                   
 dt1 <-read.csv(infile1,header=F 
          ,skip=1
            ,sep=","  
                ,quot='"' 
        , col.names=c(
                    "Pond_name",     
                    "Latitude",     
                    "Longitude",     
                    "Pond_code",     
                    "Season",     
                    "TimeOfDay",     
                    "Date",     
                    "Chlorination",     
                    "Temp_C",     
                    "CH4_uatm",     
                    "CO2_uatm",     
                    "N2O_uatm",     
                    "K_ppm",     
                    "Sr_ppm",     
                    "Mg_ppm",     
                    "Ca_ppm",     
                    "Na_ppm",     
                    "S_ppm",     
                    "Si_ppm",     
                    "Fe_ppb",     
                    "Mn_ppb",     
                    "DOC_mgL",     
                    "TDN_mgL",     
                    "Chla_ugL",     
                    "Pheo_ugL",     
                    "TP_mgL",     
                    "Cl_mgL",     
                    "SO4_mgL",     
                    "NO3_mgL",     
                    "Perc_SurfaceCover",     
                    "Perc_SubsurfaceCover",     
                    "Conductivity_uScm",     
                    "DO_mgL",     
                    "DO_perc",     
                    "pH",     
                    "Surface_m2",     
                    "volume_m3"    ), check.names=TRUE)
               
unlink(infile1)
		    
# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
                
if (class(dt1$Pond_name)!="factor") dt1$Pond_name<- as.factor(dt1$Pond_name)
if (class(dt1$Latitude)=="factor") dt1$Latitude <-as.numeric(levels(dt1$Latitude))[as.integer(dt1$Latitude) ]               
if (class(dt1$Latitude)=="character") dt1$Latitude <-as.numeric(dt1$Latitude)
if (class(dt1$Longitude)=="factor") dt1$Longitude <-as.numeric(levels(dt1$Longitude))[as.integer(dt1$Longitude) ]               
if (class(dt1$Longitude)=="character") dt1$Longitude <-as.numeric(dt1$Longitude)
if (class(dt1$Pond_code)!="factor") dt1$Pond_code<- as.factor(dt1$Pond_code)
if (class(dt1$Season)!="factor") dt1$Season<- as.factor(dt1$Season)
if (class(dt1$TimeOfDay)!="factor") dt1$TimeOfDay<- as.factor(dt1$TimeOfDay)                                   
# attempting to convert dt1$Date dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d"
tmp1Date<-as.Date(dt1$Date,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(nrow(dt1[dt1$Date != "",]) == length(tmp1Date[!is.na(tmp1Date)])){dt1$Date <- tmp1Date } else {print("Date conversion failed for dt1$Date. Please inspect the data and do the date conversion yourself.")}                                                                    
                                
if (class(dt1$Chlorination)!="factor") dt1$Chlorination<- as.factor(dt1$Chlorination)
if (class(dt1$Temp_C)=="factor") dt1$Temp_C <-as.numeric(levels(dt1$Temp_C))[as.integer(dt1$Temp_C) ]               
if (class(dt1$Temp_C)=="character") dt1$Temp_C <-as.numeric(dt1$Temp_C)
if (class(dt1$CH4_uatm)=="factor") dt1$CH4_uatm <-as.numeric(levels(dt1$CH4_uatm))[as.integer(dt1$CH4_uatm) ]               
if (class(dt1$CH4_uatm)=="character") dt1$CH4_uatm <-as.numeric(dt1$CH4_uatm)
if (class(dt1$CO2_uatm)=="factor") dt1$CO2_uatm <-as.numeric(levels(dt1$CO2_uatm))[as.integer(dt1$CO2_uatm) ]               
if (class(dt1$CO2_uatm)=="character") dt1$CO2_uatm <-as.numeric(dt1$CO2_uatm)
if (class(dt1$N2O_uatm)=="factor") dt1$N2O_uatm <-as.numeric(levels(dt1$N2O_uatm))[as.integer(dt1$N2O_uatm) ]               
if (class(dt1$N2O_uatm)=="character") dt1$N2O_uatm <-as.numeric(dt1$N2O_uatm)
if (class(dt1$K_ppm)=="factor") dt1$K_ppm <-as.numeric(levels(dt1$K_ppm))[as.integer(dt1$K_ppm) ]               
if (class(dt1$K_ppm)=="character") dt1$K_ppm <-as.numeric(dt1$K_ppm)
if (class(dt1$Sr_ppm)=="factor") dt1$Sr_ppm <-as.numeric(levels(dt1$Sr_ppm))[as.integer(dt1$Sr_ppm) ]               
if (class(dt1$Sr_ppm)=="character") dt1$Sr_ppm <-as.numeric(dt1$Sr_ppm)
if (class(dt1$Mg_ppm)=="factor") dt1$Mg_ppm <-as.numeric(levels(dt1$Mg_ppm))[as.integer(dt1$Mg_ppm) ]               
if (class(dt1$Mg_ppm)=="character") dt1$Mg_ppm <-as.numeric(dt1$Mg_ppm)
if (class(dt1$Ca_ppm)=="factor") dt1$Ca_ppm <-as.numeric(levels(dt1$Ca_ppm))[as.integer(dt1$Ca_ppm) ]               
if (class(dt1$Ca_ppm)=="character") dt1$Ca_ppm <-as.numeric(dt1$Ca_ppm)
if (class(dt1$Na_ppm)=="factor") dt1$Na_ppm <-as.numeric(levels(dt1$Na_ppm))[as.integer(dt1$Na_ppm) ]               
if (class(dt1$Na_ppm)=="character") dt1$Na_ppm <-as.numeric(dt1$Na_ppm)
if (class(dt1$S_ppm)=="factor") dt1$S_ppm <-as.numeric(levels(dt1$S_ppm))[as.integer(dt1$S_ppm) ]               
if (class(dt1$S_ppm)=="character") dt1$S_ppm <-as.numeric(dt1$S_ppm)
if (class(dt1$Si_ppm)=="factor") dt1$Si_ppm <-as.numeric(levels(dt1$Si_ppm))[as.integer(dt1$Si_ppm) ]               
if (class(dt1$Si_ppm)=="character") dt1$Si_ppm <-as.numeric(dt1$Si_ppm)
if (class(dt1$Fe_ppb)=="factor") dt1$Fe_ppb <-as.numeric(levels(dt1$Fe_ppb))[as.integer(dt1$Fe_ppb) ]               
if (class(dt1$Fe_ppb)=="character") dt1$Fe_ppb <-as.numeric(dt1$Fe_ppb)
if (class(dt1$Mn_ppb)=="factor") dt1$Mn_ppb <-as.numeric(levels(dt1$Mn_ppb))[as.integer(dt1$Mn_ppb) ]               
if (class(dt1$Mn_ppb)=="character") dt1$Mn_ppb <-as.numeric(dt1$Mn_ppb)
if (class(dt1$DOC_mgL)=="factor") dt1$DOC_mgL <-as.numeric(levels(dt1$DOC_mgL))[as.integer(dt1$DOC_mgL) ]               
if (class(dt1$DOC_mgL)=="character") dt1$DOC_mgL <-as.numeric(dt1$DOC_mgL)
if (class(dt1$TDN_mgL)=="factor") dt1$TDN_mgL <-as.numeric(levels(dt1$TDN_mgL))[as.integer(dt1$TDN_mgL) ]               
if (class(dt1$TDN_mgL)=="character") dt1$TDN_mgL <-as.numeric(dt1$TDN_mgL)
if (class(dt1$Chla_ugL)=="factor") dt1$Chla_ugL <-as.numeric(levels(dt1$Chla_ugL))[as.integer(dt1$Chla_ugL) ]               
if (class(dt1$Chla_ugL)=="character") dt1$Chla_ugL <-as.numeric(dt1$Chla_ugL)
if (class(dt1$Pheo_ugL)=="factor") dt1$Pheo_ugL <-as.numeric(levels(dt1$Pheo_ugL))[as.integer(dt1$Pheo_ugL) ]               
if (class(dt1$Pheo_ugL)=="character") dt1$Pheo_ugL <-as.numeric(dt1$Pheo_ugL)
if (class(dt1$TP_mgL)=="factor") dt1$TP_mgL <-as.numeric(levels(dt1$TP_mgL))[as.integer(dt1$TP_mgL) ]               
if (class(dt1$TP_mgL)=="character") dt1$TP_mgL <-as.numeric(dt1$TP_mgL)
if (class(dt1$Cl_mgL)=="factor") dt1$Cl_mgL <-as.numeric(levels(dt1$Cl_mgL))[as.integer(dt1$Cl_mgL) ]               
if (class(dt1$Cl_mgL)=="character") dt1$Cl_mgL <-as.numeric(dt1$Cl_mgL)
if (class(dt1$SO4_mgL)=="factor") dt1$SO4_mgL <-as.numeric(levels(dt1$SO4_mgL))[as.integer(dt1$SO4_mgL) ]               
if (class(dt1$SO4_mgL)=="character") dt1$SO4_mgL <-as.numeric(dt1$SO4_mgL)
if (class(dt1$NO3_mgL)=="factor") dt1$NO3_mgL <-as.numeric(levels(dt1$NO3_mgL))[as.integer(dt1$NO3_mgL) ]               
if (class(dt1$NO3_mgL)=="character") dt1$NO3_mgL <-as.numeric(dt1$NO3_mgL)
if (class(dt1$Perc_SurfaceCover)=="factor") dt1$Perc_SurfaceCover <-as.numeric(levels(dt1$Perc_SurfaceCover))[as.integer(dt1$Perc_SurfaceCover) ]               
if (class(dt1$Perc_SurfaceCover)=="character") dt1$Perc_SurfaceCover <-as.numeric(dt1$Perc_SurfaceCover)
if (class(dt1$Perc_SubsurfaceCover)=="factor") dt1$Perc_SubsurfaceCover <-as.numeric(levels(dt1$Perc_SubsurfaceCover))[as.integer(dt1$Perc_SubsurfaceCover) ]               
if (class(dt1$Perc_SubsurfaceCover)=="character") dt1$Perc_SubsurfaceCover <-as.numeric(dt1$Perc_SubsurfaceCover)
if (class(dt1$Conductivity_uScm)=="factor") dt1$Conductivity_uScm <-as.numeric(levels(dt1$Conductivity_uScm))[as.integer(dt1$Conductivity_uScm) ]               
if (class(dt1$Conductivity_uScm)=="character") dt1$Conductivity_uScm <-as.numeric(dt1$Conductivity_uScm)
if (class(dt1$DO_mgL)=="factor") dt1$DO_mgL <-as.numeric(levels(dt1$DO_mgL))[as.integer(dt1$DO_mgL) ]               
if (class(dt1$DO_mgL)=="character") dt1$DO_mgL <-as.numeric(dt1$DO_mgL)
if (class(dt1$DO_perc)=="factor") dt1$DO_perc <-as.numeric(levels(dt1$DO_perc))[as.integer(dt1$DO_perc) ]               
if (class(dt1$DO_perc)=="character") dt1$DO_perc <-as.numeric(dt1$DO_perc)
if (class(dt1$pH)=="factor") dt1$pH <-as.numeric(levels(dt1$pH))[as.integer(dt1$pH) ]               
if (class(dt1$pH)=="character") dt1$pH <-as.numeric(dt1$pH)
if (class(dt1$Surface_m2)=="factor") dt1$Surface_m2 <-as.numeric(levels(dt1$Surface_m2))[as.integer(dt1$Surface_m2) ]               
if (class(dt1$Surface_m2)=="character") dt1$Surface_m2 <-as.numeric(dt1$Surface_m2)
if (class(dt1$volume_m3)=="factor") dt1$volume_m3 <-as.numeric(levels(dt1$volume_m3))[as.integer(dt1$volume_m3) ]               
if (class(dt1$volume_m3)=="character") dt1$volume_m3 <-as.numeric(dt1$volume_m3)
                
# Convert Missing Values to NA for non-dates
                
dt1$CH4_uatm <- ifelse((trimws(as.character(dt1$CH4_uatm))==trimws("NA")),NA,dt1$CH4_uatm)               
suppressWarnings(dt1$CH4_uatm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$CH4_uatm))==as.character(as.numeric("NA"))),NA,dt1$CH4_uatm))
dt1$CO2_uatm <- ifelse((trimws(as.character(dt1$CO2_uatm))==trimws("NA")),NA,dt1$CO2_uatm)               
suppressWarnings(dt1$CO2_uatm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$CO2_uatm))==as.character(as.numeric("NA"))),NA,dt1$CO2_uatm))
dt1$K_ppm <- ifelse((trimws(as.character(dt1$K_ppm))==trimws("NA")),NA,dt1$K_ppm)               
suppressWarnings(dt1$K_ppm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$K_ppm))==as.character(as.numeric("NA"))),NA,dt1$K_ppm))
dt1$Sr_ppm <- ifelse((trimws(as.character(dt1$Sr_ppm))==trimws("NA")),NA,dt1$Sr_ppm)               
suppressWarnings(dt1$Sr_ppm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Sr_ppm))==as.character(as.numeric("NA"))),NA,dt1$Sr_ppm))
dt1$Mg_ppm <- ifelse((trimws(as.character(dt1$Mg_ppm))==trimws("NA")),NA,dt1$Mg_ppm)               
suppressWarnings(dt1$Mg_ppm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Mg_ppm))==as.character(as.numeric("NA"))),NA,dt1$Mg_ppm))
dt1$S_ppm <- ifelse((trimws(as.character(dt1$S_ppm))==trimws("NA")),NA,dt1$S_ppm)               
suppressWarnings(dt1$S_ppm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$S_ppm))==as.character(as.numeric("NA"))),NA,dt1$S_ppm))
dt1$Si_ppm <- ifelse((trimws(as.character(dt1$Si_ppm))==trimws("NA")),NA,dt1$Si_ppm)               
suppressWarnings(dt1$Si_ppm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Si_ppm))==as.character(as.numeric("NA"))),NA,dt1$Si_ppm))
dt1$Fe_ppb <- ifelse((trimws(as.character(dt1$Fe_ppb))==trimws("NA")),NA,dt1$Fe_ppb)               
suppressWarnings(dt1$Fe_ppb <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Fe_ppb))==as.character(as.numeric("NA"))),NA,dt1$Fe_ppb))
dt1$Mn_ppb <- ifelse((trimws(as.character(dt1$Mn_ppb))==trimws("NA")),NA,dt1$Mn_ppb)               
suppressWarnings(dt1$Mn_ppb <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Mn_ppb))==as.character(as.numeric("NA"))),NA,dt1$Mn_ppb))
dt1$Pheo_ugL <- ifelse((trimws(as.character(dt1$Pheo_ugL))==trimws("NA")),NA,dt1$Pheo_ugL)               
suppressWarnings(dt1$Pheo_ugL <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Pheo_ugL))==as.character(as.numeric("NA"))),NA,dt1$Pheo_ugL))
dt1$SO4_mgL <- ifelse((trimws(as.character(dt1$SO4_mgL))==trimws("NA")),NA,dt1$SO4_mgL)               
suppressWarnings(dt1$SO4_mgL <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$SO4_mgL))==as.character(as.numeric("NA"))),NA,dt1$SO4_mgL))
dt1$NO3_mgL <- ifelse((trimws(as.character(dt1$NO3_mgL))==trimws("NA")),NA,dt1$NO3_mgL)               
suppressWarnings(dt1$NO3_mgL <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$NO3_mgL))==as.character(as.numeric("NA"))),NA,dt1$NO3_mgL))


# Here is the structure of the input data frame:
str(dt1)                            
attach(dt1)                            

inUrl2  <- "https://pasta-s.lternet.edu/package/data/eml/edi/1927/2/1588bee09b6afab3fcd7670f839659e9" 
infile2 <- tempfile()
try(download.file(inUrl2,infile2,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile2))) download.file(inUrl2,infile2,method="auto")

                   
 dt2 <-read.csv(infile2,header=F 
          ,skip=1
            ,sep=","  
                ,quot='"' 
        , col.names=c(
                    "Pond_name",     
                    "Pond_code",     
                    "Season",     
                    "TimeOfDay",     
                    "Chlorination",     
                    "Date",     
                    "FCH4_mgCO2eqm2d1",     
                    "FCO2_mgCO2eqm2d1",     
                    "FN2O_mgCO2eqm2d1"    ), check.names=TRUE)
               
unlink(infile2)
		    
# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
                
if (class(dt2$Pond_name)!="factor") dt2$Pond_name<- as.factor(dt2$Pond_name)
if (class(dt2$Pond_code)!="factor") dt2$Pond_code<- as.factor(dt2$Pond_code)
if (class(dt2$Season)!="factor") dt2$Season<- as.factor(dt2$Season)
if (class(dt2$TimeOfDay)!="factor") dt2$TimeOfDay<- as.factor(dt2$TimeOfDay)
if (class(dt2$Chlorination)!="factor") dt2$Chlorination<- as.factor(dt2$Chlorination)                                   
# attempting to convert dt2$Date dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d"
tmp2Date<-as.Date(dt2$Date,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(nrow(dt2[dt2$Date != "",]) == length(tmp2Date[!is.na(tmp2Date)])){dt2$Date <- tmp2Date } else {print("Date conversion failed for dt2$Date. Please inspect the data and do the date conversion yourself.")}                                                                    
                                
if (class(dt2$FCH4_mgCO2eqm2d1)=="factor") dt2$FCH4_mgCO2eqm2d1 <-as.numeric(levels(dt2$FCH4_mgCO2eqm2d1))[as.integer(dt2$FCH4_mgCO2eqm2d1) ]               
if (class(dt2$FCH4_mgCO2eqm2d1)=="character") dt2$FCH4_mgCO2eqm2d1 <-as.numeric(dt2$FCH4_mgCO2eqm2d1)
if (class(dt2$FCO2_mgCO2eqm2d1)=="factor") dt2$FCO2_mgCO2eqm2d1 <-as.numeric(levels(dt2$FCO2_mgCO2eqm2d1))[as.integer(dt2$FCO2_mgCO2eqm2d1) ]               
if (class(dt2$FCO2_mgCO2eqm2d1)=="character") dt2$FCO2_mgCO2eqm2d1 <-as.numeric(dt2$FCO2_mgCO2eqm2d1)
if (class(dt2$FN2O_mgCO2eqm2d1)=="factor") dt2$FN2O_mgCO2eqm2d1 <-as.numeric(levels(dt2$FN2O_mgCO2eqm2d1))[as.integer(dt2$FN2O_mgCO2eqm2d1) ]               
if (class(dt2$FN2O_mgCO2eqm2d1)=="character") dt2$FN2O_mgCO2eqm2d1 <-as.numeric(dt2$FN2O_mgCO2eqm2d1)
                
# Convert Missing Values to NA for non-dates
                
dt2$FCO2_mgCO2eqm2d1 <- ifelse((trimws(as.character(dt2$FCO2_mgCO2eqm2d1))==trimws("NA")),NA,dt2$FCO2_mgCO2eqm2d1)               
suppressWarnings(dt2$FCO2_mgCO2eqm2d1 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt2$FCO2_mgCO2eqm2d1))==as.character(as.numeric("NA"))),NA,dt2$FCO2_mgCO2eqm2d1))


# Here is the structure of the input data frame:
str(dt2)                            
attach(dt2)      

# The analyses below are basic descriptions of the variables. After testing, they should be replaced.      
#Save as .csv
  write_csv(dt1, "rawdata/Ponds_GHG_EnvironmentalVars.csv")
  write_csv(dt2, "rawdata/CO2eq.csv")
```



# Reproduce results

## AUP features

```{r AUP library}
#library
library(tidyverse)
library(ggpubr)
library(broom)
library(errors)
library(factoextra)
library(gridExtra)
library(ggfortify)
library(knitr)
library(kableExtra)
```

```{r AUP PCA}
#Import data
  data <- read_csv("rawdata/Ponds_GHG_EnvironmentalVars.csv")
  data <- data %>%
    mutate(Chlorination = fct_relevel(Chlorination, "Yes", "No"))
  
#Select variables for PCA
  #Just day time measurements
    day <- data %>% filter(TimeOfDay == "Day")
    #Remove categorical vars and latitude and longitude
    data_PCA <- day %>% select(-c("Latitude","Longitude", "Pond_name", "Pond_code", "Season","TimeOfDay", "Date", "Chlorination"))
    
  
#When I have NAs, it is because the values were below the detection limit ~ 0
  data_PCA <- data_PCA %>% mutate(across(everything(), ~replace_na(., 0)))
  
#Transform data_PCA
  #How I have 0's in my data_PCA, I'm using an approach adding a very low concentration. In this case, I will add the half of the minimum values I get.
  minvalues <- data_PCA %>% summarise(across(where(is.numeric), ~min(.[.>0],na.rm = T)))
  data_PCA <- data_PCA %>% full_join(minvalues)
  
  data_PCA <- data_PCA  %>% mutate(across(where(is.numeric), ~case_when(. == 0 ~ last(.)/2,
                                                                        TRUE ~ .)))
  data_PCA <- data_PCA %>% mutate(across(where(is.numeric), log))
  data_PCA <- data_PCA[-length(data_PCA$Temp_C),]

#I remove dissolved oxygen because is already include in %
  data_PCA <- data_PCA %>% select(-c(DO_mgL))
  #I add depth
  data_PCA <- data_PCA %>% mutate(Depth = volume_m3/Surface_m2)
  #Remove GHG 
  data_PCA <- data_PCA %>% select(-c(CO2_uatm, CH4_uatm, N2O_uatm))
  #Rename variable for plotting PCA
  data_PCA <- data_PCA %>% rename(TP = TP_mgL,
                                  TDN = TDN_mgL,
                                  Cl = Cl_mgL,
                                  DOC = DOC_mgL,
                                  SO4 = SO4_mgL,
                                  NO3 = NO3_mgL,
                                  `Chl-a` = Chla_ugL,
                                  Pheopigments = Pheo_ugL,
                                  SurfaceVegetation = Perc_SurfaceCover,
                                  SubmergedVegetation = Perc_SubsurfaceCover,
                                  DO = DO_perc,
                                  Volume = volume_m3,
                                  Surface = Surface_m2,
                                  Temperature = Temp_C,
                                  Si = Si_ppm,
                                  Mg = Mg_ppm,
                                  Mn = Mn_ppb,
                                  S = S_ppm,
                                  Sr = Sr_ppm,
                                  Na = Na_ppm,
                                  Conductivity = Conductivity_uScm,
                                  K = K_ppm,
                                  Fe = Fe_ppb,
                                  Ca = Ca_ppm)
  #PCA
  pca_todo <- data_PCA %>% prcomp(., scale = TRUE)
```

### Figure 2

```{r}
  #Firsts two components
  plot_PC1_PC2 <- autoplot(pca_todo, x=1, y = 2, data = day, fill = "Chlorination", shape = 21, size = 3,
                           loadings = TRUE, loadings.colour = 'gray',
                           loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = "black",
                           loadings.label.repel=T,
                         loadings.label.family = "Times")+
    scale_fill_manual(values = c("#99ccff","#739900"))+
    theme_bw(base_size = 12, base_family = "Times")+
    theme(legend.position = "bottom")
  plot_PC1_PC2
  ggsave("Figure_2.pdf", width = 15, height = 15, units = "cm")
  ggsave("Figure_2.png", width = 15, height = 15, units = "cm")
```


## GHG partial pressures

```{r GHG library}
#Library
  library(tidyverse)
  library(errors)
  library(rstatix)
  library(car)
  library(emmeans)
  library(ggResidpanel)
  library(lme4)
  library(lmerTest)
```

```{r}
#Import data
data <- read_csv("rawdata/Ponds_GHG_EnvironmentalVars.csv")
data <- data %>%
    mutate(Chlorination = fct_relevel(Chlorination, "Yes", "No"))
  
  #Day data
  data <- data %>% filter(TimeOfDay == "Day")
  #Select GHGs data
  data <- data %>% select(Pond_name, Pond_code, Season, TimeOfDay, Chlorination, Date, CH4_uatm, CO2_uatm, N2O_uatm)
  
```

### Main effects
```{r GHG ANOVA, results='hide'}
##ANOVA====
  ##Assumptions====
  #- No significant outliers, normality, Homogeneity of variance
  #Identify outliers
  data_trans <- data %>% mutate(across(c(CH4_uatm, CO2_uatm, N2O_uatm), ~log(.)))
  data_trans %>% group_by(Chlorination, Season) %>% identify_outliers(CH4_uatm) %>% filter(is.extreme == "TRUE")
  data_trans %>% group_by(Chlorination, Season) %>% identify_outliers(CO2_uatm) %>% filter(is.extreme == "TRUE")
  data_trans %>% group_by(Chlorination, Season) %>% identify_outliers(N2O_uatm) %>% filter(is.extreme == "TRUE")
  
  #Checking normality
  data_trans %>% group_by(Chlorination, Season) %>% shapiro_test(CH4_uatm)
  data_trans %>% group_by(Chlorination, Season) %>% shapiro_test(CO2_uatm)
  data_trans %>% group_by(Chlorination, Season) %>% shapiro_test(N2O_uatm)
  
  #Check homogeneity
  data_trans %>%
    group_by(Season) %>%
    levene_test(CH4_uatm ~ Chlorination)
  data_trans %>%
    group_by(Season) %>%
    levene_test(CO2_uatm ~ Chlorination)
  data_trans %>%
    group_by(Season) %>%
    levene_test(N2O_uatm ~ Chlorination)

##Perform the ANOVA====  
  ##ANOVA TWO-FACTORS repeated measurements
  data_trans <- data_trans %>% mutate(Pond_name = as.factor(Pond_name))
  ##CH4
  res.aov_CH4 <- anova_test(data = data_trans, dv = CH4_uatm, wid = Pond_name, between = Chlorination, within = Season)
  get_anova_table(res.aov_CH4)
  #CO2
  res.aov_CO2 <- anova_test(data = data_trans, dv = CO2_uatm, wid = Pond_name, between = Chlorination, within = Season)
  get_anova_table(res.aov_CO2)
  
##Non parametric test for N2O====
  #Linear mixed models
  data_trans <- data_trans %>% mutate(Chlorination = as.factor(Chlorination), Season = as.factor(Season))
  
  #N2O  
  model <- lmer(N2O_uatm ~ Chlorination * Season + (1 | Pond_name), data = data_trans)
  resid_panel(model)
  summary(model)
  anova(model, ddf = "Kenward-Roger")
  Posthoc_N2O <- emmeans(model, specs = c("Chlorination", "Season"))
  intervals <- contrast(Posthoc_N2O, method = "pairwise", adjust = "tukey")
  intervals_pairwise_N2O <- confint(intervals)
  Posthoc_N2O <- pairs(Posthoc_N2O, adjust = "tukey") %>% as_data_frame()
  

```

#### ANOVA tables

```{r ANOVA table}
  print("CO2")
  print(res.aov_CO2)
  print("CH4")
  print(res.aov_CH4)
  print("N2O")
  print(anova(model, ddf = "Kenward-Roger"))
```

### Posthoc test
```{r GHG Posthoc}
##Post-hoc tests====
    #TUKEY
  Pond_name <- factor(data_trans$Pond_name)
  Season <- factor(data_trans$Season)
  Chlorination <- factor(data_trans$Chlorination)
  CH4_uatm <- data_trans$CH4_uatm
  tukey_CH4 <- tukey_hsd(res.aov_CH4, CH4_uatm ~ Season*Chlorination)
  CO2_uatm <- data_trans$CO2_uatm
  tukey_CO2 <- tukey_hsd(res.aov_CO2, CO2_uatm ~ Season*Chlorination)

  #Export
  print("CH4")
  print(tukey_CH4)
  print("N2O")
  print(intervals_pairwise_N2O)
```
  
### Figure 3
```{r GHG plot}
###==========Letters for plotting===========================
  library(multcompView)
  ##CO2
  tukey_df <- tukey_CO2[3:8,] %>%
    select(group1, group2, p.adj) %>%
    arrange(p.adj)
  
  pvalues <- tukey_df$p.adj
  names(pvalues) <- paste(tukey_df$group1, tukey_df$group2, sep = "-")
  
  # Letters for comparison
  letters <- multcompLetters(pvalues)
  letters_CO2 <- data.frame(
    Combination = names(letters$Letters), 
    Letters = letters$Letters,
    variable = "CO2_uatm"
  )
  
  ##CH4
  tukey_df <- tukey_CH4[3:8,] %>%
    select(group1, group2, p.adj) %>%
    arrange(p.adj)
  
  pvalues <- tukey_df$p.adj
  names(pvalues) <- paste(tukey_df$group1, tukey_df$group2, sep = "-")
  
  #Letters
  letters <- multcompLetters(pvalues)
  letters_CH4 <- data.frame(
    Combination = names(letters$Letters), 
    Letters = letters$Letters,
    variable = "CH4_uatm"
  )
  
  ##N2O
  pvalues <- Posthoc_N2O$p.value
  factores <- Posthoc_N2O$contrast
  factores <- str_split_fixed(factores, " - ", 2) %>% as.data.frame()
  factores <-factores %>% separate_wider_delim(V1, delim = " ", names = c("V1_Chlorination", "V1_Season"))
  factores <-factores %>% separate_wider_delim(V2, delim = " ", names = c("V2_Chlorination", "V2_Season"))
  
  #Similar format than CO2 and CH4
  factores <- factores %>% mutate(Reconstruido = paste(paste(V1_Season, V1_Chlorination, sep = ":"), paste(V2_Season, V2_Chlorination, sep = ":"), sep = "-"))
  
  names(pvalues) <- factores$Reconstruido
  
  #Letters
  letters <- multcompLetters(pvalues)
  letters_N2O <- data.frame(
    Combination = names(letters$Letters), 
    Letters = letters$Letters,
    variable = "N2O_uatm"
  )  
  #Join all letter df
  letters <- letters_CO2 %>% bind_rows(letters_CH4) %>% bind_rows(letters_N2O)
  
  #Df letter and combinations
  data_comb <- data %>%
    mutate(Combination = paste(Season, Chlorination, sep = ":")) 
  
##Join letter for plotting
  #long format
  data_l <- data_comb %>%  pivot_longer(-c(Pond_name, Pond_code, Season, TimeOfDay, Chlorination, Date, Combination), names_to = "variable", values_to = "uatm")
  
  #Join with letters
  data_l <- data_l %>% left_join(letters)
  
  #I add atmosphere references values
  data_l <- data_l %>% mutate(Atm_ref = case_when(variable == "CH4_uatm" ~ 1.92026,
                                                  variable == "CO2_uatm" ~ 424,
                                                  variable == "N2O_uatm" ~ 0.33653))
  #Label position en y
  data_l <- data_l %>% group_by(Season, variable) %>% 
    mutate(label_y = max(uatm, na.rm = T)) %>% 
    ungroup()
  
  #reorder variables
  data_l <- data_l %>% mutate(variable = fct_relevel(variable, "CO2_uatm"))
  
##-------------PLOT - LOG SCALE Y-AXIS-----------------
  ggplot(data_l, aes(x = Season, y = uatm, fill = Chlorination))+
    geom_violin(trim = F)+
    geom_text(aes(label = Letters, y = label_y), position = position_dodge(width = 0.8), vjust = -5, family = "Times") +
    geom_jitter(alpha = 0.4,position=position_jitterdodge(jitter.width = 0.2, dodge.width = 0.9), shape = 21)+
    geom_hline(aes(yintercept = Atm_ref, linetype = "Atmosphere ref (NOAA, 2023)"))+
    scale_linetype_manual(values = 2)+
    scale_fill_manual(values = c("#99ccff","#739900"))+
    scale_y_log10(expand = expansion(mult = c(0.1, 0.2)), labels = scales::label_number(accuracy = 1, big.mark = ","), guide = "axis_logticks")+
    labs(x = "Chlorination", y = expression(paste("Partial pressure (", mu, "atm)")), linetype = NULL)+
    facet_wrap(~variable, scales = "free", labeller = as_labeller(c(CH4_uatm = "CH[4]", CO2_uatm = "CO[2]", N2O_uatm = "N[2]*O"), default = label_parsed))+
    theme_bw(base_size = 12, base_family = "Times")+
    theme(legend.position = "bottom", text = element_text(family = "Times"))
  ggsave("Figure_3.pdf", width = 16, height = 10, units = "cm")
  ggsave("Figure_3.png", width = 16, height = 10, units = "cm")
```

### Summary

```{r}
##Summary
  data_l %>% drop_errors() %>% group_by(variable) %>% 
    summarise(Mean = mean(uatm, na.rm = T),
              Sd = sd(uatm, na.rm = T),
              Median = median(uatm, na.rm = T),
              Q1 = quantile(uatm, 0.25, na.rm = T),
              Q3 = quantile(uatm, 0.75, na.rm = T),
              Min = min(uatm, na.rm = T),
              Max = max(uatm, na.rm = T),
              SE = sd(uatm, na.rm = T)/sqrt(sum(!is.na(uatm))), #Not considering NAs
              n = sum(!is.na(uatm)))%>%
  mutate(across(where(is.numeric), ~ round(.x, 2))) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE,
                position = "center") %>%
  scroll_box(width = "100%", height = "400px")
  
  #Per group
  data_l %>% drop_errors() %>% group_by(variable, Chlorination, Season) %>%
    summarise(Mean = mean(uatm, na.rm = T),
              Sd = sd(uatm, na.rm = T),
              Median = median(uatm, na.rm = T),
              Q1 = quantile(uatm, 0.25, na.rm = T),
              Q3 = quantile(uatm, 0.75, na.rm = T),
              Min = min(uatm, na.rm = T),
              Max = max(uatm, na.rm = T),
              SE = sd(uatm, na.rm = T)/sqrt(sum(!is.na(uatm))), #Not considering NAs
              n = sum(!is.na(uatm)))%>%
  mutate(across(where(is.numeric), ~ round(.x, 2))) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE,
                position = "center") %>%
  scroll_box(width = "100%", height = "400px")
```


## Diel changes

```{r}
###Library====
library(tidyverse)
library(errors)
library(ggpubr)
library(rstatix)
library(car)
library(lmerTest)
library(ggpubr)

#Import data
data <- read_csv("rawdata/Ponds_GHG_EnvironmentalVars.csv")
data <- data %>%
    mutate(Chlorination = fct_relevel(Chlorination, "Yes", "No"))

#Ponds sampled at night
night <- data %>% filter(TimeOfDay == "Night")
#Select all information for these ponds
data <- data %>% filter(Pond_code %in% night$Pond_code)

#GHGs data
data <- data %>% select(Pond_name, Pond_code, Season, TimeOfDay, Chlorination, Date, CH4_uatm, CO2_uatm, N2O_uatm)


#Log transformation before perform the analyses
data_trans <- data %>% mutate(across(c(CH4_uatm, CO2_uatm, N2O_uatm), ~log(.)))
data_trans <- data_trans %>% mutate(Pond_name = as.factor(Pond_name))
data_trans <- drop_errors(data_trans) %>% ungroup()

##Assumptions====
#- Outliers, normality, Homogeneity of variance
#Identify outliers
data_trans <- data %>% mutate(across(c(CH4_uatm, CO2_uatm, N2O_uatm), ~log(.)))
data_trans %>% group_by(Chlorination, TimeOfDay) %>% identify_outliers(CH4_uatm) %>% filter(is.extrem = FALSE)
data_trans %>% group_by(Chlorination, TimeOfDay) %>% identify_outliers(CO2_uatm) %>% filter(is.extrem = FALSE)
data_trans %>% group_by(Chlorination, TimeOfDay) %>% identify_outliers(N2O_uatm) %>% filter(is.extrem = FALSE)

#Checking normality
data_trans %>% group_by(Chlorination, Season, TimeOfDay) %>% shapiro_test(CH4_uatm)
data_trans %>% group_by(Chlorination, Season, TimeOfDay) %>% shapiro_test(CO2_uatm)
data_trans %>% group_by(Chlorination, Season, TimeOfDay) %>% shapiro_test(N2O_uatm)

#Check homogeneity
data_trans %>%
  group_by(Season, Chlorination) %>%
  levene_test(CH4_uatm ~ TimeOfDay)
data_trans %>%
  group_by(Season, Chlorination) %>%
  levene_test(CO2_uatm ~ TimeOfDay)
data_trans %>%
  group_by(Season, Chlorination) %>%
  levene_test(N2O_uatm ~ TimeOfDay)


##Transform
  data_trans_l <- data_trans %>% select(Season, TimeOfDay, Chlorination, CH4_uatm, CO2_uatm, N2O_uatm) %>% 
    pivot_longer(-c(Season, TimeOfDay, Chlorination), names_to = "Species")
  
  #T_test
  options(width = 200)
  #CO2
  data_trans_l %>% filter(Species == "CO2_uatm") %>%
    group_by(Chlorination, Season) %>% 
    t_test(value ~ TimeOfDay, paired = T)%>%
    adjust_pvalue(method = "holm")
  #CH4
  data_trans_l %>% filter(Species == "CH4_uatm") %>%
    group_by(Chlorination, Season) %>% 
    t_test(value ~ TimeOfDay, paired = T)%>%
    adjust_pvalue(method = "holm")
  #N2O
  data_trans_l %>% filter(Species == "N2O_uatm") %>%
    group_by(Chlorination, Season) %>% 
    t_test(value ~ TimeOfDay, paired = T)%>%
    adjust_pvalue(method = "holm")

  ##-------------LOG SCALE Y-AXIS-----------------
  data_l <- data %>% pivot_longer(-c(Pond_name, Pond_code, Season, TimeOfDay, Chlorination, Date), names_to = "variable", values_to = "uatm")
  #Description plots
  #I add atmosphere references values
  data_l <- data_l %>% mutate(Atm_ref = case_when(variable == "CH4_uatm" ~ 1.92026,
                                                  variable == "CO2_uatm" ~ 424,
                                                  variable == "N2O_uatm" ~ 0.33653))
  #reorder variables
  data_l <- data_l %>% mutate(variable = fct_relevel(variable, "CO2_uatm"))
  my_comparisons <- list(c("Day", "Night"))
  
```

### Figure 4

```{r plot 4, fig.height=7.87, fig.width=5.91}
ggplot(data_l, aes(x = Chlorination, y = uatm, fill = interaction(Chlorination, TimeOfDay, sep = "-", lex.order = T)))+
  geom_violin(trim = F)+
  #geom_text(aes(label = Letters, y = label_y), position = position_dodge(width = 0.8), vjust = -2.5) +
  #stat_compare_means(method = "t.test", label = "p.signif", comparisons = my_comparisons)+
  geom_pwc(aes(group = TimeOfDay), tip.length = 0.02,
           method = "t_test", label = "p.signif", bracket.nudge.y = 0.3, p.adjust.method = "holm", family = "Times")+
  geom_jitter(alpha = 0.4,position=position_jitterdodge(jitter.width = 0.2, dodge.width = 0.9))+
  geom_hline(aes(yintercept = Atm_ref, linetype = "Atmosphere ref (NOAA, 2023)"))+
  scale_linetype_manual(values = 2)+
  scale_fill_manual(values=c("#E0F0FF","#99ccff", "#BFD47F","#739900"))+
  scale_alpha_manual(values = c(1, 0.2))+ 
  scale_y_log10(expand = expansion(mult = c(0.1, 0.12)), labels = scales::label_number(accuracy = 1, big.mark = ","), guide = "axis_logticks")+
  labs(x = "Chlorination", y = expression(paste("Partial pressure (", mu, "atm)")), linetype = NULL, fill = "Chlorination - Time of Day", color = NULL)+
  facet_grid(variable~Season, scales = "free_y", labeller = labeller(variable = as_labeller(c(CH4_uatm = "CH[4]", CO2_uatm = "CO[2]", N2O_uatm = "N[2]*O"), default = label_parsed)))+
  theme_bw(base_size = 12, base_family = "Times")+
  guides(fill = guide_legend(title.position = "top", 
                        title.hjust = 0.5))+
  theme(legend.position = "bottom", legend.text=element_text(size=10),
        legend.title = element_text(size = 10, face = "bold"), legend.key.size = unit(0.7, 'lines'),
        legend.box = "vertical",
        text = element_text(family = "Times"),
        strip.text.y = element_text(angle = 0))+
  guides(color = "none")
  ggsave("Figure_4.pdf", width = 16, height = 20, units = "cm") 
  ggsave("Figure_4.png", width = 16, height = 20, units = "cm") 
```


## Drivers

```{r Drivers, results='hide'}
#library
  library(tidyverse)
  library(car)
  library(corrplot)
  library(GGally)
  library(errors)
  library(fuzzySim)

  #Import data
  data <- read_csv("rawdata/Ponds_GHG_EnvironmentalVars.csv")
  data <- data %>%
    mutate(Chlorination = fct_relevel(Chlorination, "Yes", "No"))
  #Day data
  data <- data %>% filter(TimeOfDay == "Day")
  
  #When I have NAs they are actually 0 (or close to 0) because we couldn't detect it. So:
  data <- data %>%
    mutate(across(where(is.numeric) & !all_of(c("CO2_uatm", "CH4_uatm", "N2O_uatm")), ~replace_na(., 0)))
  
  #Remove DO mgL because it is in %
  data <- data %>% select(-c(DO_mgL, Latitude, Longitude))
  
  #Reduce correlated variables
  data <- data %>% select(-c(Mg_ppm, S_ppm, Sr_ppm, Ca_ppm, Si_ppm, Na_ppm, TDN_mgL))
  data <- data %>% mutate(Depth = volume_m3/Surface_m2) %>% 
    select(-c(Surface_m2, volume_m3))
  
  ##Residual distribution
  Data_l <- data %>% select(-c(Date)) %>%
    pivot_longer(-c(TimeOfDay, Pond_name, Pond_code, Chlorination, Season), names_to =  "variable")
  ggplot(Data_l, aes(sample = value))+
    stat_qq()+
    stat_qq_line()+
    facet_wrap(~variable, scales = "free")
  #Log
  ggplot(Data_l, aes(sample = log(value)))+
    stat_qq()+
    stat_qq_line()+
    facet_wrap(~variable, scales = "free")
  
  #Transform data 0's area handle adding a constant. In this case, it will be added the half of the minimum values register.
  data <- data %>% drop_errors()
  minvalues <- data %>% summarise(across(where(is.numeric), ~min(.[.>0],na.rm = T)))
  data <- data %>% full_join(minvalues)
  
  data <- data  %>% mutate(across(where(is.numeric), ~case_when(. == 0 ~ last(.)/2,
                                                                TRUE ~ .)))
  data <- data %>% mutate(across(where(is.numeric), log))
  data <- data[-length(data$Pond_name),]
  
  #Explore collinearity
  #Select just numerical variables
  data_num <- data %>% select(-c(TimeOfDay, Pond_name, Pond_code, Chlorination, Season, Date))
  
  #Inf as NAs
  cor_matrix <- cor(data_num, use = "pairwise.complete.obs")
  corrplot(cor_matrix, type = "upper", order = "hclust", 
           tl.col = "black", tl.srt = 45, sig.level = 0.01, insig = "blank")
  #Just high correlated values > 0.7
  cor_matrix2 <- cor_matrix
  cor_matrix2[cor_matrix2 > -0.7 & cor_matrix2 < 0.7] <- 0
  corrplot(cor_matrix2, type = "upper", order = "hclust", 
           tl.col = "black", tl.srt = 45, sig.level = 0.01, insig = "blank")
  #Reduce collinearity
  data <- data %>% select(-c(Conductivity_uScm, DOC_mgL))
  
  #For CH4
  #Step wise regression
  CH4 <- data %>% select(-c(CO2_uatm, N2O_uatm, Pond_name,Pond_code, Season, TimeOfDay, Chlorination, Date))
  CH4 <- drop_errors(CH4)
  any(is.na(CH4))
  sapply(CH4, function(col) any(is.infinite(col)))
  modelCH4 <- lm(CH4_uatm ~ ., data = CH4)

  modelCH4 <- step(modelCH4, direction = "both")  
  
  modelCH4
  summary(modelCH4)
  vif(modelCH4)
  
  #For CO2
  #Step wise regression
  CO2 <- data %>% select(-c(CH4_uatm, N2O_uatm, Pond_name,Pond_code, Season, TimeOfDay, Chlorination, Date))
  
  modelCO2 <- lm(CO2_uatm ~ ., data = CO2)

  modelCO2 <- step(modelCO2, direction = "both")  
  
  modelCO2
  summary(modelCO2)
  vif(modelCO2)
  modelCO2$anova
  
  #For N2O
  #Step wise regression
  N2O <- data %>% select(-c(CO2_uatm, CH4_uatm, Pond_name,Pond_code, Season, TimeOfDay, Chlorination, Date))
  
  modelN2O <- lm(N2O_uatm ~ ., data = N2O)
  
  modelN2O <- step(modelN2O, direction = "both")  
  
  modelN2O
  summary(modelN2O)
  vif(modelN2O)

##Remove non significants variables  
  
  #CH4
  modelCH4_trimmed <- modelTrim(modelCH4) 
  modelCH4_trimmed
  summary(modelCH4_trimmed)
  AIC(modelCH4_trimmed)
  vif(modelCH4_trimmed)
  resid_panel(modelCH4_trimmed)
  
  #CO2
  modelCO2_trimmed <- modelTrim(modelCO2)
  modelCO2_trimmed
  summary(modelCO2_trimmed)
  AIC(modelCO2_trimmed)
  vif(modelCO2_trimmed)
  resid_panel(modelCO2_trimmed)
  
  #N2O
  modelN2O_trimmed <- modelTrim(modelN2O)
  modelN2O_trimmed
  summary(modelN2O_trimmed)
  AIC(modelN2O_trimmed)
  vif(modelN2O_trimmed) 
  resid_panel(modelN2O_trimmed)
```

```{r export resid plots, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
resid_panel(modelCO2_trimmed)
ggsave("Figure_S1.pdf", width = 16, height = 15, units = "cm")
ggsave("Figure_S1.png", width = 16, height = 15, units = "cm")
resid_panel(modelCH4_trimmed)
ggsave("Figure_S2.pdf", width = 16, height = 15, units = "cm")
ggsave("Figure_S2.png", width = 16, height = 15, units = "cm")
resid_panel(modelN2O_trimmed)
ggsave("Figure_S3.pdf", width = 16, height = 15, units = "cm")
ggsave("Figure_S3.png", width = 16, height = 15, units = "cm")
```


```{r models}
  print("CO2")
  summary(modelCO2_trimmed)
  AIC(modelCO2_trimmed)
  print("CH4")
  summary(modelCH4_trimmed)
  AIC(modelCH4_trimmed)
  print("N2O")
  summary(modelN2O_trimmed)
  AIC(modelN2O_trimmed)
  
  
  #Create the table
    #define function to extract overall p-value of model
    overall_p <- function(my_model) {
      f <- summary(my_model)$fstatistic
      p <- unname(pf(f[1],f[2],f[3],lower.tail=F))
      attributes(p) <- NULL
      return(p)
    }
  #Table
  Table_models <-data.frame(Model = as.character(c(formula(modelCO2_trimmed), formula(modelCH4_trimmed),formula(modelN2O_trimmed))),
             DF = c(df.residual(modelCO2_trimmed), df.residual(modelCH4_trimmed),df.residual(modelN2O_trimmed)),
             RMSE = c(round(sigma(modelCO2_trimmed), 2), round(sigma(modelCH4_trimmed), 2),round(sigma(modelN2O_trimmed), 2)),
             R2 = c(round(summary(modelCO2_trimmed)$r.squared, 2), round(summary(modelCH4_trimmed)$r.squared, 2), round(summary(modelN2O_trimmed)$r.squared, 2)),
             R2adj = c(round(summary(modelCO2_trimmed)$adj.r.squared, 2), round(summary(modelCH4_trimmed)$adj.r.squared, 2), round(summary(modelN2O_trimmed)$adj.r.squared, 2)),
             Pvalue = c(format.pval(overall_p(modelCO2_trimmed)), format.pval(overall_p(modelCH4_trimmed)), format.pval(overall_p(modelN2O_trimmed))),
             R2adj = c(round(summary(modelCO2_trimmed)$fstatistic[1], 2), round(summary(modelCH4_trimmed)$fstatistic[1], 2), round(summary(modelN2O_trimmed)$fstatistic[1], 2)),
             AIC = c(round(AIC(modelCO2_trimmed), 2), round(AIC(modelCH4_trimmed), 2),round(AIC(modelN2O_trimmed), 2)))
  
```

### Table 1 - Model


```{r, Table models}
  Table_models %>%
  mutate(across(where(is.numeric), ~ round(.x, 2))) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE,
                position = "center")
  
```


## Difussive flux

```{r}
 library(tidyverse)
  library(errors)
  library(rstatix)
  library(ggResidpanel)
  library(lmerTest)

#Import data
  data <- read_csv("rawdata/CO2eq.csv")
  data <- data %>%
  mutate(Chlorination = fct_relevel(Chlorination, "Yes", "No"))

  #long format
  data_l <- data %>%  pivot_longer(-c(Pond_name, Pond_code, Season, TimeOfDay, Chlorination, Date), names_to = "variable", values_to = "Flux_CO2eq")

  #reorder variables
  data_l <- data_l %>% mutate(variable = fct_relevel(variable, "FCO2_mgCO2eqm2d1"))
```

### Figure 5

```{r}
  #Plot
  #Mean between summer and winter
  data_l <- data_l %>% group_by(Chlorination, variable, Pond_name) %>% 
    summarise(Flux_CO2eq = mean(Flux_CO2eq, na.rm = T)) %>% 
    ungroup()
  add_total_flux <- data_l %>% group_by(Chlorination, Pond_name) %>% summarise(Flux_CO2eq = sum(Flux_CO2eq)) %>% 
  ungroup()  %>% 
  mutate(variable = "Total")

data_l %>% bind_rows(add_total_flux) %>% 
  mutate(variable = fct_relevel(variable, "FCO2_mgCO2eqm2d1")) %>% 
  ggplot(aes(x = Chlorination, y = Flux_CO2eq, fill = variable))+
  geom_boxplot()+
  stat_summary(fun = mean, geom = "point", shape = 8, size = 2, color = "black", position = position_dodge(width = 0.75))+
  geom_hline(yintercept = 0)+
  #geom_jitter(alpha = 0.2, position=position_jitterdodge(), shape = 21)+
  scale_fill_manual(values = c("gray40","#e6b800","#cc2900", "gray90"), labels = c(expression(CO[2]), expression(CH[4]), expression(N[2]*O, "Total")))+
  labs(y = expression(paste("GHG flux (mg CO"[2]*" eq · m"^{-2}*"· d"^{-1}, ")")), fill = NULL)+
  theme_bw(base_size = 12, base_family = "Times")  
  ggsave("Figure_5.pdf", width = 16, height = 15, units = "cm") 
  ggsave("Figure_5.png", width = 16, height = 15, units = "cm")
```

### Wilcoxon tests
```{r, results='hide'}
#Mean per pond (winter and summer))  
  # data_l <- data_l %>% group_by(Pond_name, Chlorination, variable) %>% summarise(Flux_CO2eq = mean(Flux_CO2eq)) %>% 
  #   ungroup()
  #Transform data
  CH4 <- data_l %>% filter(variable == "FCH4_mgCO2eqm2d1") %>% mutate(Flux_CO2eq = log(Flux_CO2eq))
  #Check outlier, normality and homogeneity
  CH4 %>% group_by(Chlorination, variable) %>% identify_outliers(Flux_CO2eq) %>% filter(is.extreme == "TRUE")
  CH4 %>% group_by(Chlorination) %>% shapiro_test(Flux_CO2eq)
  CH4 %>% levene_test(Flux_CO2eq ~ Chlorination)
  ##Wilcoxon (U man whitney)
  res.wilcox_CH4 <- wilcox_test(data = CH4, Flux_CO2eq ~ Chlorination)
  
  
#CO2  
  #Transform data
  CO2 <- data_l %>% filter(variable == "FCO2_mgCO2eqm2d1") %>% mutate(Flux_CO2eq = log(abs(Flux_CO2eq))*sign(Flux_CO2eq))
  #Check outlier, normality and homogeneity
  CO2 %>% group_by(Chlorination, variable) %>% identify_outliers(Flux_CO2eq)%>% filter(is.extreme == "TRUE")
  CO2 %>% group_by(Chlorination) %>% shapiro_test(Flux_CO2eq)
  CO2 %>% drop_errors()%>% levene_test(Flux_CO2eq ~ Chlorination)
  
  #Wilcoxon
  res.wilcox_CO2 <- wilcox_test(data = CO2, Flux_CO2eq ~ Chlorination)
  res.wilcox_CO2

#N2O
  #Transform data
  N2O <-  data_l %>% filter(variable == "FN2O_mgCO2eqm2d1") %>% mutate(Flux_CO2eq = log(Flux_CO2eq))
  #Check outlier, normality and homogeneity
  N2O %>% group_by(Chlorination, variable) %>% identify_outliers(Flux_CO2eq)%>% filter(is.extreme == "TRUE")
  N2O %>% group_by(Chlorination) %>% shapiro_test(Flux_CO2eq)
  N2O %>% levene_test(Flux_CO2eq ~ Chlorination)
  
  #Wilcoxon
  res.wilcox_N2O <- wilcox_test(data = N2O, Flux_CO2eq ~ Chlorination)
  res.wilcox_N2O
```

#### Per gas species
```{r}
#CH4
res.wilcox_CH4
#CO2
res.wilcox_CO2
#N2O
res.wilcox_N2O
```

#### Total CO2 equivalent
```{r}
##Compare total CO2eq flux (CO2+CH4+N2O)
  #long format
  # data_l <- data %>%  pivot_longer(-c(Pond_name, Pond_code, Season, TimeOfDay, Chlorination, Date), names_to = "variable", values_to = "Flux_CO2eq")
  ##Total CO2eq
  data_total_l <- data_l %>% group_by(Chlorination, Pond_name) %>% summarise(Total_FCO2eq = sum(Flux_CO2eq)) %>% 
    ungroup()  
  #Annual mean
  data_total_l <- data_total_l %>% 
    mutate(Log_Total_FCO2eq = log(Total_FCO2eq)) %>% 
    ungroup()
  
  #Check outlier, normality and homogeneity
  data_total_l %>% group_by(Chlorination) %>% identify_outliers(Log_Total_FCO2eq)%>% filter(is.extreme == "TRUE")
  data_total_l %>% group_by(Chlorination) %>% shapiro_test(Log_Total_FCO2eq)
  data_total_l %>% drop_errors() %>% levene_test(Log_Total_FCO2eq ~ Chlorination)
  

  ##Wilcoxon (U man whitney)
  res.wilcox_TF <- wilcox_test(data = data_total_l, Log_Total_FCO2eq ~ Chlorination)
  res.wilcox_TF
  
```

### Summary
```{r }
##Summary
  data_total_l %>% group_by(Chlorination) %>% 
    summarise(Median = median(Total_FCO2eq, na.rm = T),
              Min = min(Total_FCO2eq, na.rm = T),
              Max = max(Total_FCO2eq, na.rm = T),
              Q1 = quantile(Total_FCO2eq, na.rm = T, probs = 0.25),
              Q3 = quantile(Total_FCO2eq, na.rm = T, probs = 0.75),
              IQR = IQR(Total_FCO2eq, na.rm = T),
              Mean = mean(Total_FCO2eq, na.rm = T))
  data_l %>% group_by(variable, Chlorination) %>% 
    summarise(Median = median(Flux_CO2eq, na.rm = T),
              Min = min(Flux_CO2eq, na.rm = T),
              Max = max(Flux_CO2eq, na.rm = T),
              Q1 = quantile(Flux_CO2eq, na.rm = T, probs = 0.25),
              Q3 = quantile(Flux_CO2eq, na.rm = T, probs = 0.75),
              IQR = IQR(Flux_CO2eq, na.rm = T),
              Mean = mean(Flux_CO2eq, na.rm = T))
  
  #% calculate per pond and summarise by Chlorination
  Contribution <- data_l %>% left_join(select(data_total_l, - Log_Total_FCO2eq)) %>% 
    rowwise() %>% 
    mutate(Per_contribution = Flux_CO2eq/Total_FCO2eq*100) %>% 
    ungroup()
  
  Contribution %>% group_by(variable, Chlorination) %>% 
    summarise(Median = median(Per_contribution, na.rm = T),
              Min = min(Per_contribution, na.rm = T),
              Max = max(Per_contribution, na.rm = T),
              Q1 = quantile(Per_contribution, na.rm = T, probs = 0.25),
              Q3 = quantile(Per_contribution, na.rm = T, probs = 0.75),
              IQR = IQR(Per_contribution, na.rm = T),
              Mean = mean(Per_contribution, na.rm = T))

##Calculation of the relevance of this GHGs emission
  #The per capita emission in Barcelona in 2022 was ~ 2.1 t CO2eq 
  #The peak emission in BCN was reached in 2005 with 5·10⁶ t CO2eq, and in 2022 it was 3.5 · 10⁶ t CO2eq 
  #Info: https://bcnroc.ajuntament.barcelona.cat/jspui/bitstream/11703/133640/1/Balanc%cc%a7_Energia_2022.pdf
  
  #The total surface area of all artificial water bodies in BCN is ~ 130000 m².
  
  #The average emission of a tourist is 11.6 kg CO2 eq/day #https://www-sciencedirect-com.sire.ub.edu/science/article/pii/S0261517718302206
  
  #Median
  median(data_total_l$Total_FCO2eq*132000*365/1000000000, na.rm = T) #t CO2eq
  #Q1
  quantile(data_total_l$Total_FCO2eq*132000*365/1000000000, na.rm = T, probs = 0.25)
  #Q3
  quantile(data_total_l$Total_FCO2eq*132000*365/1000000000, na.rm = T, probs = 0.75)
  #It is equivalent to:
  median(data_total_l$Total_FCO2eq*132000*365/1000000000, na.rm = T)/2.1 #inhabitants

##Print results  
  print("CO2")
  res.wilcox_CO2
  print("CH4")
  res.wilcox_CH4
  print("N2O")
  res.wilcox_N2O
  print("Total")
  res.wilcox_TF
  print("Summary")
  data_total_l %>% group_by(Chlorination) %>% 
    summarise(Median = median(Total_FCO2eq, na.rm = T),
              Min = min(Total_FCO2eq, na.rm = T),
              Max = max(Total_FCO2eq, na.rm = T),
              Q1 = quantile(Total_FCO2eq, na.rm = T, probs = 0.25),
              Q3 = quantile(Total_FCO2eq, na.rm = T, probs = 0.75),
              IQR = IQR(Total_FCO2eq, na.rm = T))
  data_l %>% group_by(variable, Chlorination) %>% 
    summarise(Median = median(Flux_CO2eq, na.rm = T),
              Min = min(Flux_CO2eq, na.rm = T),
              Max = max(Flux_CO2eq, na.rm = T),
              Q1 = quantile(Flux_CO2eq, na.rm = T, probs = 0.25),
              Q3 = quantile(Flux_CO2eq, na.rm = T, probs = 0.75),
              IQR = IQR(Flux_CO2eq, na.rm = T))
  print("ContributionPercentage")
  Contribution %>% group_by(variable, Chlorination) %>% 
    summarise(Median = median(Per_contribution, na.rm = T),
              Min = min(Per_contribution, na.rm = T),
              Max = max(Per_contribution, na.rm = T),
              Q1 = quantile(Per_contribution, na.rm = T, probs = 0.25),
              Q3 = quantile(Per_contribution, na.rm = T, probs = 0.75),
              IQR = IQR(Per_contribution, na.rm = T))
  #Median
  median(data_total_l$Total_FCO2eq*132000*365/1000000000, na.rm = T) #t CO2eq
  #Q1
  quantile(data_total_l$Total_FCO2eq*132000*365/1000000000, na.rm = T, probs = 0.25)
  #Q3
  quantile(data_total_l$Total_FCO2eq*132000*365/1000000000, na.rm = T, probs = 0.75)
  #It is equivalent to:
  median(data_total_l$Total_FCO2eq*132000*365/1000000000, na.rm = T)/2.1 #inhabitants
```


# Suplementary material

```{r Supplementary material, results='hide'}

##Libraries----
  library(tidyverse)
  library(lmerTest)
  library(lme4)
  library(errors)
  library(ggResidpanel)
  library(rstatix)

#Import data
data <- read_csv("rawdata/Ponds_GHG_EnvironmentalVars.csv")
data <- data %>%
    mutate(Chlorination = fct_relevel(Chlorination, "Yes", "No"))

#Day data
data <- data %>% filter(TimeOfDay == "Day")
  
  #I add depth variable
  data <- data %>% mutate(Depth = volume_m3/Surface_m2)
  #I only take it into account for the percentages, which are values of 0, and I add 0.0001 to them
  data <- data %>% mutate(across(c(Perc_SubsurfaceCover, Perc_SurfaceCover), ~case_when(. == 0 ~ 0.01,
                                                                                        TRUE ~ .)))
  data <- data %>% mutate(across(where(is.numeric), ~case_when(. == 0 ~ NA,
                                                                                     TRUE ~ .)))
  #Select variable included in the table
  data_comparison <- data %>% select(Pond_name, Season, Chlorination, DO_mgL, DO_perc, Conductivity_uScm, pH, Temp_C,
                                     Chla_ugL, Pheo_ugL, Perc_SubsurfaceCover, Perc_SurfaceCover,
                                     DOC_mgL, TDN_mgL, TP_mgL,
                                     Cl_mgL, SO4_mgL, NO3_mgL,
                                     K_ppm, Mg_ppm, Ca_ppm, Ca_ppm, Na_ppm, S_ppm, Si_ppm,
                                     Fe_ppb, Mn_ppb, Sr_ppm,
                                     volume_m3, Surface_m2, Depth)
  
##Transform data----
  data_trans <- data_comparison

  data_trans <- data_trans %>% mutate(across(where(is.numeric), log))

##Longformat data
  Data_comparison_l <- data_trans %>%
    pivot_longer(-c(Pond_name, Chlorination, Season), names_to =  "Var")
  
##Check assumptions----
 #I'm going to apply linear mixed models so the assumption should be check it after model fitting
  
##t-studen or U Mann Whitney or lmer----
  #I use a loop 
  Vars <- unique(Data_comparison_l$Var)
  models <- list()
  for(i in Vars){
    print(i)
    #i <- Vars[26]
    data_model <- Data_comparison_l %>% filter(Var == i) %>% drop_na()
    model <- lmer(value ~ Chlorination + (1 | Pond_name) + (1 | Season), data = data_model)
    models[[length(models) + 1]] <- model
  }
  names(models) <- Vars
  anovas_models <- lapply(models, anova)
  anovas_models
  #Extract p values
  pvalues <- lapply(anovas_models, function(x){ return(x$`Pr(>F)`)})
  pvalues <- do.call(rbind, pvalues) %>% as.data.frame() %>% 
    rownames_to_column(var = "Var") %>% 
    rename(pvalue = V1)

##Check models residuals====
  #Extract residuals
  residuos <- lapply(models, residuals)
  residuos <- do.call(cbind, residuos) %>% as.data.frame() %>%
    rownames_to_column(var = "rowname")
  residuos_l <- residuos %>% pivot_longer(-rowname, names_to = "Variable", values_to  = "Residuals")
  #Extract fitted values
  fitted_values <- lapply(models, fitted)
  fitted_values <- do.call(cbind, fitted_values) %>% as.data.frame() %>%
    rownames_to_column(var = "rowname")
  fitted_values_l <- fitted_values %>% pivot_longer(-rowname, names_to = "Variable", values_to  = "Fitted")
  data_check <- left_join(residuos_l, fitted_values_l)
  
  data_table <- data %>% select(Pond_name, Season, Chlorination, DO_mgL, DO_perc, Conductivity_uScm, pH, Temp_C,
                                Chla_ugL, Pheo_ugL, Perc_SubsurfaceCover, Perc_SurfaceCover,
                                DOC_mgL, TDN_mgL, TP_mgL,
                                Cl_mgL, SO4_mgL, NO3_mgL,
                                K_ppm, Mg_ppm, Ca_ppm, Ca_ppm, Na_ppm, S_ppm, Si_ppm, 
                                Fe_ppb, Mn_ppb, Sr_ppm,
                                volume_m3, Surface_m2, Depth)
  
  Data_l <- data_table %>%
    pivot_longer(-c(Pond_name, Chlorination, Season), names_to =  "Var")
  ##Surface and volume=====
        #For surface and volume i can't use repeatd data, i select just one season and apply t.test
        #Select just surface and volume
        data_SV <- Data_l %>% filter(Season == "Summer") %>% filter(Var %in% c("Surface_m2", "volume_m3"))
        Data_l <- Data_l %>% filter(!Var %in% c("Surface_m2", "volume_m3")) %>% bind_rows(data_SV)
        #Transform data to meet assumptions
        data_SV <- data_SV %>% mutate(value = log(value))
        #Check extreme outliers
        data_SV %>% group_by(Chlorination, Var) %>% identify_outliers(value) %>% filter(is.extreme == "TRUE")
        #Check normality
        data_SV %>% group_by(Chlorination, Var) %>% shapiro_test(value)
        #Check homogeneity
        data_SV %>%
          group_by(Var) %>%
          levene_test(value ~ Chlorination)
        #Now t.test
        pvalue_SV <- data_SV %>%  group_by(Var) %>% t_test(value ~ Chlorination) %>% rename(pvalue = p) %>% 
          select(Var, pvalue)
        #Join to the others pvalues
        pvalues <- pvalues %>% filter(!Var %in% c("Surface_m2", "volume_m3")) %>% bind_rows(pvalue_SV)

        #For depth I have to apply a lmm
        data_D <- Data_l %>% filter(Season == "Summer") %>% filter(Var %in% c("Depth"))
        Data_l <- Data_l %>% filter(!Var %in% c("Depth")) %>% bind_rows(data_D)
        #Transform data to meet assumptions
        data_D <- data_D %>% mutate(value = log(value))
        #Check extreme outliers
        data_D %>% group_by(Chlorination, Var) %>% identify_outliers(value) %>% filter(is.extreme == "TRUE")
        #Check normality
        data_D %>% group_by(Chlorination, Var) %>% shapiro_test(value)
        #Check homogeneity
        data_D %>%
          group_by(Var) %>%
          levene_test(value ~ Chlorination)
        #Now t.test
        pvalue_D <- data_D %>%  group_by(Var) %>% wilcox_test(value ~ Chlorination) %>% rename(pvalue = p) %>% 
          select(Var, pvalue)
        #Join to the others pvalues
        pvalues <- pvalues %>% filter(!Var %in% c("Depth")) %>% bind_rows(pvalue_D)
        
  
  Tabla_Results <- Data_l %>% group_by(Chlorination, Var) %>% 
    summarise(Mean = mean(value, na.rm = T),
              Sd = sd(value, na.rm = T),
              Median = median(value, na.rm = T),
              Q1 = quantile(value, 0.25, na.rm = T),
              Q3 = quantile(value, 0.75, na.rm = T),
              Min = min(value, na.rm = T),
              Max = max(value, na.rm = T),
              SE = sd(value, na.rm = T)/sqrt(sum(!is.na(value))), #Not considering NAs
              n = sum(!is.na(value))) %>% 
    ungroup()
  Tabla_Results <- Tabla_Results %>% left_join(pvalues)
  
  
  #Assign categories
  Physical <- c("Temp", "Surface_m2", "volume_m3", "Depth")
  Biological <- c("Chla_ugL", "Pheo_ugL", "Perc_SurfaceCover", "Perc_SubsurfaceCover")
  Chemical <- Tabla_Results %>% select(Var) %>% filter(!Var %in% Physical & !Var %in% Biological) %>% pull()
  Tabla_Results <- Tabla_Results %>% mutate(Category = case_when(Var %in% Physical ~ "Physical",
                                                 Var %in% Biological ~ "Biological",
                                                 Var %in% Chemical ~ "Chemical"))
  
  #Arrange by category and reorder by column
  Tabla_Results <- Tabla_Results %>% arrange(Chlorination, Category) %>% 
    relocate(Chlorination, Category, Var, pvalue, Mean, Sd, Median, Q1, Q3, Min, Max)
  
```

## Table S1
```{r}


  Tabla_Results %>%
  mutate(across(where(is.numeric), ~ round(.x, 2))) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE,
                position = "center") %>%
  scroll_box(width = "100%", height = "400px")

```


```

